---
title: "Letter-winged Kite neuromorphology"
author: "Aubrey Keirnan, Vera Weisbecker and Andrew Iwaniuk"
date: "August 17th, 2020"
output: html_document
---

# load libraries required
```{r}

library(ape)         # CRAN v5.5 
library(nlme)        # CRAN v3.1-153
library(geiger)      # CRAN v2.0.7
library(rstudioapi)  # CRAN v0.13
library(devtools)    # CRAN v2.4.3
library(phytools)    # CRAN v0.7-90
library(phylolm)     # CRAN v2.6.2
library(performance) # CRAN v0.8.0

#Note that for my installations, I have to specify the path to the library on C:/

#install_github("JeroenSmaers/evomap", lib = "C:/Program Files/R/R-4.0.4/library" )
#install_github("khabbazian/l1ou", lib = "C:/Program Files/R/R-4.0.4/library" )

library(l1ou)        # [github::khabbazian/l1ou] v1.42
library (evomap)    #0.0.0.9000



#this sets the working directory to the home path, which doesn't happen on everyone's computers
setwd(dirname(getActiveDocumentContext()$path))

load(file = "../Data/Processed/LWK_data_processed.rda")

```

#prepare analyses that ask if Elanus is different - phylo ANOVA and prediction intervals

```{r}


#Pick out the right species, check

tree<-consensus_tree
plot(tree,label.offset = 0.25); tiplabels()
All<-1:length(tree$tip.label)
Elanus<-getTips(tree,findMRCA(tree,c("Elanus_axillaris","Elanus_scriptus", "Elanus_caeruleus")))

grpI_01<-rep("XXX",length(rownames(data)))
    grpI_01[All]<-"1"
    grpI_01[Elanus]<-"2"
    grpI_01<-as.factor(grpI_01); names(grpI_01)<-rownames(data)
    summary(grpI_01); length(grpI_01); length(names(summary(grpI_01)))
    tree$tip.label[which(grpI_01=="XXX")]
        tipCol<-rep("black",length(tree$tip.label))
        tipCol[which(grpI_01=="1")]<-"red"
        tipCol[which(grpI_01=="2")]<-"darkgreen"
        #Check taht the right species were chosen
        plot(tree,tip.color = tipCol)
    
dataTemp<-pruneSample(Av_data,tree,setdiff(All,Elanus))$dataPruned
treeTemp<-pruneSample(Av_data,tree,setdiff(All,Elanus))$treePruned

name.check(treeTemp,dataTemp)

```

#PGLS models and Evomap confidence/prediction intervals. The CI and PI code is later used for making plot lines. All data are logged because the expectation is that they evolve to scale exponentially.

#NOTE here and in figure it has to be endocast SA minus optic lobe SA

```{r}



#Orbit to Occ_Hinge_mm
#The below also computing PIs (depends on prep code above) because we can expect the Elanus to be different
lm_orbD <- lm(log (orbD) ~ log (Occ_Hinge_mm),  data = Av_data)

perf_orbD <- check_model(lm_orbD)

gls_orbD <- gls( log (orbD) ~ log (Occ_Hinge_mm), correlation = corBrownian(1,consensus_tree, form = ~Species), data = Av_data)
summary(gls_orbD)
#Evomap confidence intervals.  PI code needs emphasizing that the continuous variables are numeric!
OrbD_gm_CI <- gls.ci(log (Av_data$orbD), log (Av_data$Occ_Hinge_mm),vcv(consensus_tree))
OrbD_gm_PI <- gls.pi(log (as.numeric(dataTemp$orbD)), log (as.numeric(dataTemp$Occ_Hinge_mm)), vcv(treeTemp),3)
        


#Optic Foramen to Occ_Hinge_mm

lm_optF <- lm(log(optF) ~  log(Occ_Hinge_mm), data = Av_data )

perf_optF <- check_model(lm_optF)

gls_optF <- gls( log(optF) ~  log(Occ_Hinge_mm), correlation = corBrownian(1,consensus_tree, form = ~Species), data = Av_data)
summary(gls_optF)
#Evomap confidence intervals
OptF_gm_CI <- gls.ci(log (Av_data$optF), log (Av_data$Occ_Hinge_mm),vcv(consensus_tree))
OptF_gm_PI <- gls.pi(log (as.numeric(dataTemp$optF)), log (as.numeric(dataTemp$Occ_Hinge_mm)), vcv(treeTemp),3)


#Optic Foramen to Orbit. 

lm_optF_orbD <- lm( log (optF) ~ log (orbD), data = Av_data)

perf_optF_orbD <- check_model(lm_optF_orbD)

gls_optF_orbD <- gls( log (optF) ~ log (orbD), correlation = corBrownian(1,consensus_tree, form = ~Species), data = Av_data)
summary(gls_optF_orbD)
#Evomap confidence intervals
OptF_OrbD_gm_CI <- gls.ci(log (Av_data$optF), log (Av_data$orbD),vcv(consensus_tree))
OptF_OrbD_gm_PI <- gls.pi(log (as.numeric(dataTemp$optF)), log (as.numeric(dataTemp$orbD)), vcv(treeTemp),3)





#Surface areas . This is using confidence intervals because from the scatterplot we have no reason to assume that Elanus is different.

#First, doing the linear model underlying the gls to see if the model meets conditions of linearity, homogeneity of variance, outliers, residual normality using the performance package

lm_optLSA <- lm(log(optLSA) ~ log(Occ_Hinge_mm),data = Av_data)

perf_optLSA <- check_model(lm_optLSA)

#Then running gls

gls_optLSA <- gls(  log(optLSA) ~ log(Occ_Hinge_mm), correlation = corBrownian(1,consensus_tree, form= ~Species), data = Av_data)
summary(gls_optLSA)

#Evomap confidence intervals
OptLSA_gm_CI <- gls.ci(log(Av_data$optLSA), log(Av_data$Occ_Hinge_mm),vcv(consensus_tree))




#Endocast to Geometric mean

lm_EndoLSA <- lm( log(BrSA_mmsquared) ~ log(Occ_Hinge_mm), data = Av_data)

perf_EndoLSA <- check_model(lm_optLSA)

gls_EndoLSA <- gls(  log(BrSA_mmsquared) ~ log(Occ_Hinge_mm), correlation = corBrownian(1,consensus_tree, form= ~Species), data = Av_data)
summary(gls_EndoLSA)

#Evomap confidence intervals
EndoLSA_gm_CI <- gls.ci(log(Av_data$BrSA_mmsquared) , log(Av_data$Occ_Hinge_mm),vcv(consensus_tree))



#Optic Lobes to Endocast

lm_optLSA_endo <- lm ( log(optLSA) ~ log( BrSA_mmsquared - optLSA), data = Av_data)

perf_optLSA_endo <- check_model(lm_optLSA_endo)

gls_optLSA_endo <- gls( log(optLSA) ~ log( BrSA_mmsquared - optLSA), correlation = corBrownian(1,consensus_tree, form=~Species), data = Av_data)
summary(gls_optLSA_endo)

#Evomap confidence intervals
OptLSA_Endo_gm_CI <- gls.ci( log(Av_data$optLSA),  log(Av_data$BrSA_mmsquared-Av_data$optLSA),vcv(consensus_tree))




#subtracting Wulst from endo surface area here but it makes zero difference.


lm_Wulst_Endo <- lm(log(Wulst_mmsquared) ~ log(BrSA_mmsquared-Wulst_mmsquared),data = Av_data)

perf_Wulst_Endo <- check_model(lm_Wulst_Endo)


gls_Wulst_Endo <- gls( log(Wulst_mmsquared) ~ log(BrSA_mmsquared-Wulst_mmsquared), correlation = corBrownian(1,consensus_tree, form = ~Species), data = Av_data)
summary(gls_Wulst_Endo)
#Evomap confidence intervals
Wulst_Endo_CI <- gls.ci(log(Av_data$Wulst_mmsquared),log( Av_data$BrSA_mmsquared-Av_data$Wulst_mmsquared),vcv(consensus_tree))




```




#PGLS results table. Sooo clunkily coded but easier than manually typing in. rbind allows easy addition and swapping (as opposed to hard-coding row numbers).
```{r}

#Set up a table with the relevant column names

PGLS_results_hinge <- data.frame("Dep" = integer(0), "Pred"= integer(0), "t"= integer(0), "p"= integer(0))

#Start it off with row 1 

PGLS_results_hinge [1,] <-   c( "log Orb. Diam.", "log Occ. Hinge", round (summary(gls_orbD)$tTable[2,3:4],3))


PGLS_results_hinge  <-  rbind(PGLS_results_hinge, c( "log Opt. For.", "log Occ. Hinge",  round (summary(gls_optF)$tTable[2,3:4],3)))


PGLS_results_hinge <- rbind(PGLS_results_hinge, c("log Opt. L. Surf", "log Occ. Hinge", round(summary(gls_optLSA)$tTable[2,3:4],3)))



PGLS_results_hinge <-rbind (PGLS_results_hinge, c("log Endoc. surf.", "log Occ. Hinge", round (summary(gls_EndoLSA)$tTable[2,3:4],3)))



#The csv for these is written below with all the normality tests



```



#pANCOVAS - code by Jeroen Smaers

#Do Elanines have larger orbits, smaller Optic foramina for their size; and do they have smaller optic foramina relative to orbit size?
#This code comparesa  model where there is no intercept difference between Elanines with a model where there is one, but the slope is assumed to be the same. Significant p-values mean that the model where Elanines have a different intercept is significantly better.

```{r}


#Orbit diameter vs.  Occ. Hinge

Y<-"orbD"; X<-"Occ_Hinge_mm";

#Note, data logged here
data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")

Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_01_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)

#model is ok

lm_orbD_Occ_Hinge_mm_Elanus <- lm(Dependent~grpI_01 + Independent,data)


perf_OrbD_Occ_Hinge_mm_Elanus <- check_model(lm_orbD_Occ_Hinge_mm_Elanus )


Sigma<-vcv(tree)
OrbD_Occ_Hinge_mm_pANOVA <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)


#Optic foramen diameter vs geom mean

Y<-"optF"; X<-"Occ_Hinge_mm";

data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")

Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_optFGM_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)

lm__optF_Occ_Hinge_mm_Elanus <- lm(Dependent~grpI_01 + Independent,data)

perf_optF_Occ_Hinge_mm_Elanus <- check_model(lm__optF_Occ_Hinge_mm_Elanus )


Sigma<-vcv(tree)
OptF_Occ_Hinge_mm_pANOVA <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)


#For CIRCUS optic lobe

#Pick out the right species, check

tree<-consensus_tree
plot(tree,label.offset = 0.25); tiplabels()
All<-1:length(tree$tip.label)
Circus <- 4

grpI_01<-rep("XXX",length(rownames(data)))
    grpI_01[All]<-"1"
    grpI_01[Circus]<-"2"
    grpI_01<-as.factor(grpI_01); names(grpI_01)<-rownames(data)
    summary(grpI_01); length(grpI_01); length(names(summary(grpI_01)))
    tree$tip.label[which(grpI_01=="XXX")]
        tipCol<-rep("black",length(tree$tip.label))
        tipCol[which(grpI_01=="1")]<-"red"
        tipCol[which(grpI_01=="2")]<-"darkgreen"
        plot(tree,tip.color = tipCol)
    
#Optic lobe vs geomean

Y<-"optLSA"; X<-"Occ_Hinge_mm";

data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")

lm_optLSA_Occ_hinge_Circus <- lm(Dependent~grpI_01 + Independent,data)
perf_optLSA_Occ_hinge_Circus <- check_model(lm_optLSA_geomean_Circus)

Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_01_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)

Sigma<-vcv(tree)
OptL_Occ_Hinge_mm_pANOVA_Circus <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)

pANOVA_summary <- rbind(OrbD_Occ_Hinge_mm_pANOVA, OptF_Occ_Hinge_mm_pANOVA,OptL_Occ_Hinge_mm_pANOVA_Circus)

write.csv(pANOVA_summary, file = "../Tables/pANOVA_output_hinge.csv")

```




#Some checks for model performance

```{r}

#A pdf with model linearity, collinearity,and residual normality ; not reported, but just for pretty checks. Takes a long time so commented out

# pdf(file = "../Figures/performance_plots_hinge.pdf", height = 10, width = 8, onefile = TRUE)
# 
# perf_optLSA
# perf_EndoLSA
# 
# perf_orbD
# perf_optF
# 
# perf_OrbD_Occ_Hinge_mm_Elanus
# perf_optF_Occ_Hinge_mm_Elanus
# 
# 
# dev.off()



normchecks <- c("lm_orbD","lm_optF","lm_optLSA","lm_EndoLSA","lm_orbD_Occ_Hinge_mm_Elanus",  "lm__optF_Occ_Hinge_mm_Elanus", "lm_optLSA_Occ_hinge_Circus")

model_res_norm <- matrix(nrow = length(normchecks), ncol = 2)

for (i in c(1: length(normchecks))){
  
  
  temp <- check_normality(get(normchecks[i]))
  
  model_res_norm [i,1] <- paste (normchecks[i])
  model_res_norm[i,2] <- round( temp[1], digits = 3)
  
  
}




PGLS_results <- cbind(PGLS_results_hinge, model_res_norm[c(1:nrow(PGLS_results_hinge)),2])


write.csv(PGLS_results, file = "../Tables/PGLS_output_hinge.csv")

```



##############Plots- set up

```{r, echo=TRUE}

colsSpecies=c("red3","firebrick4", "indianred1", "forestgreen", "hotpink",  "blue", "darkblue","darkslateblue", "dark orange", "lightsalmon4", "wheat3","deepskyblue", "azure4", "skyblue", "cyan",   "blueviolet")

#adding some plotting parameters that mean you can change them all at once for each subplot. You can also specify colours this way
lwd_abline <- 2
lwd_CI <- 2
lty_CI <- 2
col_CI <- "hotpink"

names(colsSpecies)<- levels (as.factor(Av_data$GenSpec))
pchPLOT=c(19,8,8, 17, 15)
cexes = c(2,2.5,2.5,2,2)
Av_data$desig <- rep(1,length(rownames(Av_data))); Av_data$desig[which(Av_data$species=="scriptus")] <- 2;
Av_data$desig[which(Av_data$species=="axillaris")] <- 3;

Av_data$desig[which(Av_data$species=="caeruleus")] <- 3



Full_data$desig <- rep(1,length(rownames(Full_data)));
Full_data$desig[c(which(Full_data$species=="scriptus"),which(Full_data$species=="axillaris"), which(Full_data$species == "caeruleus"))] <- 2
Full_data$desig[which(Full_data$species=="subcristata")] <- 4;
Full_data$desig[which(Full_data$species=="assimilis")] <- 5



```


#This is plotting the evomap CI intervals based confidence intervals. Note that I am using upper/lower 2.5% of the distribution (a total of 5% because we're looking at both tails).


#Orbit/optic foramen plots. Really lazy in double-plotting but whatever ;-); saves me setting up a complicated line of plot just to prepare


```{r}

pdf("../Figures/Orbit_optic_hinge.pdf", height = 10, width = 5)

par(mfrow=c(4,1), bty="n", mar = c(rep(4,4)))

xlim1 <-  c(3.15,3.85)  
ylim_orb_geo <- c(2.65,3.4)
ylim_opt_geo <- c(0.8,1.5)
ylim_opt_orb <- c(0.8,1.5)

cexesOrbopt<-cexes-0.6

#Orbit diameter against Occ_Hinge_mm
plot(log (Full_data$orbD)~ log (Full_data$Occ_Hinge_mm),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Orbit Diameter (mm)", xlab="log Occ. Hinge (mm)", xlim=xlim1, ylim = ylim_orb_geo, bty="o")


#Here we make the grey PI polygon. To make that appear behind the plot, I have to repeat the plot command. Not elegant but works!

polygon(c(OrbD_gm_PI$PI.plot$X, rev (OrbD_gm_PI$PI.plot$X)), c(OrbD_gm_PI$PI.plot$Lower2.5, rev(OrbD_gm_PI$PI.plot$Upper2.5)), col = grey (0.9), border = NA )

par(new=TRUE)

plot(log (Full_data$orbD)~ log (Full_data$Occ_Hinge_mm),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Orbit Diameter (mm)", xlab="log Occ. Hinge (mm)", xlim=xlim1, ylim = ylim_orb_geo,bty="o");

lines(OrbD_gm_CI$CI.plot$X,OrbD_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
  lines(OrbD_gm_CI$CI.plot$X,OrbD_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)

  #I am here using the abline from the evomap code, but the pgls with brownian settings results in teh same line
abline (OrbD_gm_CI$model[1,1], OrbD_gm_CI$model[2,1])

mtext(text="A",side=1,line=-10,outer=FALSE, adj = 0.04,cex=0.7)

#Optic foramen diameter against Occ_Hinge_mm
plot(log (Full_data$optF)~ log (Full_data$Occ_Hinge_mm),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Optic Foramen Diameter (mm)", xlab="log Occ. Hinge (mm)",xlim=xlim1, ylim=ylim_opt_geo,bty="o");
text((Full_data$optF)~(Full_data$Occ_Hinge_mm), labels='', adj=1, cex=0.7)


polygon(c(OptF_gm_PI$PI.plot$X, rev (OptF_gm_PI$PI.plot$X)), c(OptF_gm_PI$PI.plot$Lower2.5, rev(OptF_gm_PI$PI.plot$Upper2.5)), col = grey (0.9), border = NA )

par(new=TRUE)


plot(log (Full_data$optF)~ log (Full_data$Occ_Hinge_mm),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Optic Foramen Diameter (mm)", xlab="log Occ. Hinge (mm)",xlim=xlim1, ylim=ylim_opt_geo,bty="o");
text((Full_data$optF)~(Full_data$Occ_Hinge_mm), labels='', adj=1, cex=0.7)

mtext(text="B",side=1,line=-10,outer=FALSE, adj = 0.04,cex=0.7)

abline (OptF_gm_CI$model[1,1], OptF_gm_CI$model[2,1],lwd = lwd_abline)
#The coefficients from the pgls are [1] intercept, [2] slope, and you can fit this to the graph like yay - there are cosmetics you can look up by typing ?abline.
#abline( gls_optF$coefficients[1],gls_optF$coefficients[2], lwd = lwd_abline  )

lines(OptF_gm_CI$CI.plot$X,OptF_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
lines(OptF_gm_CI$CI.plot$X,OptF_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)

  
     

plot(1, type = "n",                         # Remove all elements of plot
     xlim = c(0, 10), ylim = c(0, 10),
     axes=F,
     xlab=NA, ylab=NA
     
     )
legend(-0.6,11 , levels(as.factor(Full_data$GenSpec)) , text.font = 3, pch= c(19,19,19,17,15,8,8,8,rep(19,8)), col=colsSpecies[levels(as.factor(Full_data$GenSpec))], cex=1.1, pt.cex=1.4, bty="n",ncol=3 )

dev.off()
  
```

#Endocast surface/size, optic lobe/size,

```{r}

pdf("../Figures/Surfaces_hinge.pdf", height = 10, width = 10)

#The layout and legend plotting is inspired by code by Emma Sherratt :-) in Weisbecker et al. 2021.

#This tells it to make 3 rows and (therefore 2 columns) with images 1-4 each occupying one position and image 5 occupying 2 positions. Thanks to Emma Sherratt for this trick
mat <- matrix(c(1,2,3,3), nrow=2,byrow=T)

# set the size of fields
layout(mat, widths=c(1,1,0.5,0.5), heights=c(0.3,0.3,0.12))
#Plot PC1 vs PC2
par(mar=c(4, 5, 0.5, 1), cex.lab = 1.5, cex.axis =1.5)  # sets the margins c(bottom, left, top, right)



  
#Optic lobe against Occ_Hinge_mm
plot(log(Full_data$optLSA)~log(Full_data$Occ_Hinge_mm),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexes[Full_data$desig], ylab="log Av. Optic Lobe Surf. Area (mm2)", xlab=" log Occ. Hinge");


#The coefficients from the pgls are [1] intercept, [2] slope, and you can fit this to the graph like yay - there are cosmetics you can look up by typing ?abline.
abline( gls_optLSA$coefficients[1],gls_optLSA$coefficients[2], lwd = lwd_abline  )

  lines(OptLSA_gm_CI$CI.plot$X,OptLSA_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
  lines(OptLSA_gm_CI$CI.plot$X,OptLSA_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)

  mtext(text="A",side=1,line=-25,outer=FALSE, adj = 0.04, cex=0.9)

  
#Endocast surf. lobe against Occ_Hinge_mm
plot(log (Full_data$BrSA_mmsquared) ~ log (Full_data$Occ_Hinge_mm),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexes[Full_data$desig], ylab= "log Endoc. Surf. Area (mm2)", xlab="log Occ. Hinge");

abline( gls_EndoLSA$coefficients[1],gls_EndoLSA$coefficients[2], lwd = lwd_abline  )

  lines(EndoLSA_gm_CI$CI.plot$X,EndoLSA_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
  lines(EndoLSA_gm_CI$CI.plot$X,EndoLSA_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)
  
  mtext(text="B",side=1,line=-25,outer=FALSE, adj = 0.04, cex=0.9)

  
  #Make new empty plot for legend
  
  plot(1, type = "n",                         # Remove all elements of plot
     xlim = c(0, 10), ylim = c(0, 10),
     axes=F,
     xlab=NA, ylab=NA
     
     )
legend(0,10 , levels(as.factor(Full_data$GenSpec)) , text.font = 3, pch= c(19,19,19,17,15,8,8,8,rep(19,8)), col=colsSpecies[levels(as.factor(Full_data$GenSpec))], cex=1.5, pt.cex=2.1, bty="n",ncol=4, y.intersp = 1.2 )
  
dev.off()
  
```

