---
title: "Analysis and figure code for for: Not like night and day: the nocturnal Letter-winged Kite does not differ from diurnal congeners in orbit or endocast morphology"
author: "Vera Weisbecker, Jeroen Smaers,Aubrey Keirnan, and Andrew Iwaniuk"
date: "February 3rd, 2022"
output: html_document
---

# load libraries required
```{r}

library(ape)         # CRAN v5.5 
library(nlme)        # CRAN v3.1-153
library(geiger)      # CRAN v2.0.7
library(rstudioapi)  # CRAN v0.13
library(devtools)    # CRAN v2.4.3
library(phytools)    # CRAN v0.7-90
library(performance) # CRAN v0.8.0

#Note that for my installations, I have to specify the path to the library on C:/

#install_github("JeroenSmaers/evomap", lib = "C:/Program Files/R/R-4.0.4/library" )
#install_github("khabbazian/l1ou", lib = "C:/Program Files/R/R-4.0.4/library" )

library(l1ou)        # [github::khabbazian/l1ou] v1.42
library (evomap)    #0.0.0.9000



#this sets the working directory to the home path, which doesn't happen on everyone's computers
setwd(dirname(getActiveDocumentContext()$path))

load(file = "../Data/Processed/LWK_data_processed.rda")

```

#prepare analyses that ask if Elanus is different - phylo ANOVA and prediction intervals

```{r}


#Pick out the right species, check

tree<-consensus_tree
plot(tree,label.offset = 0.25); tiplabels()
All<-1:length(tree$tip.label)
Elanus<-getTips(tree,findMRCA(tree,c("Elanus_axillaris","Elanus_scriptus", "Elanus_caeruleus")))

grpI_01<-rep("XXX",length(rownames(data)))
    grpI_01[All]<-"1"
    grpI_01[Elanus]<-"2"
    grpI_01<-as.factor(grpI_01); names(grpI_01)<-rownames(data)
    summary(grpI_01); length(grpI_01); length(names(summary(grpI_01)))
    tree$tip.label[which(grpI_01=="XXX")]
        tipCol<-rep("black",length(tree$tip.label))
        tipCol[which(grpI_01=="1")]<-"red"
        tipCol[which(grpI_01=="2")]<-"darkgreen"
        #Check taht the right species were chosen
        plot(tree,tip.color = tipCol)
    
dataTemp<-pruneSample(Av_data,tree,setdiff(All,Elanus))$dataPruned
treeTemp<-pruneSample(Av_data,tree,setdiff(All,Elanus))$treePruned

name.check(treeTemp,dataTemp)

```

# endocast surface areas with and without blood vessels and trigeminal lobes are exchangeable
```{r}

plot(Full_data$BrSA_mmsquared_withBloodv_Trig~Full_data$BrSA_mmsquared)

summary(lm(Full_data$BrSA_mmsquared_withBloodv_Trig~Full_data$BrSA_mmsquared))

#text(Full_data$BrSA_mmsquared_withBloodv_Trig~Full_data$BrSA_mmsquared, labels=Full_data$GenSpec )



```



#How large are differences between the 2 collared sparrowhawks?
```{r}

A_cirro <- Data_raw[which(Data_raw$species == "cirrocephalus"),]

min (A_cirro[2,6:16]*100/A_cirro[1,6:16])
max (A_cirro[2,6:16]*100/A_cirro[1,6:16])

```

#PGLS models and Evomap confidence/prediction intervals. The CI and PI code is later used for making plot lines. All data are logged because the expectation is that they evolve to scale exponentially.

#NOTE here and in figure it has to be endocast SA minus optic lobe SA

```{r}



#Orbit to Geomean
#The below also computing PIs (depends on prep code above) because we can expect the Elanus to be different
lm_orbD <- lm(log (orbD) ~ log (geomean),  data = Av_data)

perf_orbD <- check_model(lm_orbD)

gls_orbD <- gls( log (orbD) ~ log (geomean), correlation = corBrownian(1,consensus_tree, form = ~Species), data = Av_data)
summary(gls_orbD)
#Evomap confidence intervals.  PI code needs emphasizing that the continuous variables are numeric!
OrbD_gm_CI <- gls.ci(log (Av_data$orbD), log (Av_data$geomean),vcv(consensus_tree))
OrbD_gm_PI <- gls.pi(log (as.numeric(dataTemp$orbD)), log (as.numeric(dataTemp$geomean)), vcv(treeTemp),3)
        


#Optic Foramen to Geomean

lm_optF <- lm(log(optF) ~  log(geomean), data = Av_data )

perf_optF <- check_model(lm_optF)

gls_optF <- gls( log(optF) ~  log(geomean), correlation = corBrownian(1,consensus_tree, form = ~Species), data = Av_data)
summary(gls_optF)
#Evomap confidence intervals
OptF_gm_CI <- gls.ci(log (Av_data$optF), log (Av_data$geomean),vcv(consensus_tree))
OptF_gm_PI <- gls.pi(log (as.numeric(dataTemp$optF)), log (as.numeric(dataTemp$geomean)), vcv(treeTemp),3)


#Optic Foramen to Orbit. 

lm_optF_orbD <- lm( log (optF) ~ log (orbD), data = Av_data)

perf_optF_orbD <- check_model(lm_optF_orbD)

gls_optF_orbD <- gls( log (optF) ~ log (orbD), correlation = corBrownian(1,consensus_tree, form = ~Species), data = Av_data)
summary(gls_optF_orbD)
#Evomap confidence intervals
OptF_OrbD_gm_CI <- gls.ci(log (Av_data$optF), log (Av_data$orbD),vcv(consensus_tree))
OptF_OrbD_gm_PI <- gls.pi(log (as.numeric(dataTemp$optF)), log (as.numeric(dataTemp$orbD)), vcv(treeTemp),3)





#Surface areas . This is using confidence intervals because from the scatterplot we have no reason to assume that Elanus is different.

#First, doing the linear model underlying the gls to see if the model meets conditions of linearity, homogeneity of variance, outliers, residual normality using the performance package

lm_optLSA <- lm(log(optLSA) ~ log(geomean),data = Av_data)

perf_optLSA <- check_model(lm_optLSA)

#Then running gls

gls_optLSA <- gls(  log(optLSA) ~ log(geomean), correlation = corBrownian(1,consensus_tree, form= ~Species), data = Av_data)
summary(gls_optLSA)

#Evomap confidence intervals
OptLSA_gm_CI <- gls.ci(log(Av_data$optLSA), log(Av_data$geomean),vcv(consensus_tree))




#Endocast to Geometric mean

lm_EndoLSA <- lm( log(BrSA_mmsquared) ~ log(geomean), data = Av_data)

perf_EndoLSA <- check_model(lm_optLSA)

gls_EndoLSA <- gls(  log(BrSA_mmsquared) ~ log(geomean), correlation = corBrownian(1,consensus_tree, form= ~Species), data = Av_data)
summary(gls_EndoLSA)

#Evomap confidence intervals
EndoLSA_gm_CI <- gls.ci(log(Av_data$BrSA_mmsquared) , log(Av_data$geomean),vcv(consensus_tree))



#Optic Lobes to Endocast

lm_optLSA_endo <- lm ( log(optLSA) ~ log( BrSA_mmsquared - optLSA), data = Av_data)

perf_optLSA_endo <- check_model(lm_optLSA_endo)

gls_optLSA_endo <- gls( log(optLSA) ~ log( BrSA_mmsquared - optLSA), correlation = corBrownian(1,consensus_tree, form=~Species), data = Av_data)
summary(gls_optLSA_endo)

#Evomap confidence intervals
OptLSA_Endo_gm_CI <- gls.ci( log(Av_data$optLSA),  log(Av_data$BrSA_mmsquared-Av_data$optLSA),vcv(consensus_tree))




#subtracting Wulst from endo surface area here but it makes zero difference.


lm_Wulst_Endo <- lm(log(Wulst_mmsquared) ~ log(BrSA_mmsquared-Wulst_mmsquared),data = Av_data)

perf_Wulst_Endo <- check_model(lm_Wulst_Endo)


gls_Wulst_Endo <- gls( log(Wulst_mmsquared) ~ log(BrSA_mmsquared-Wulst_mmsquared), correlation = corBrownian(1,consensus_tree, form = ~Species), data = Av_data)
summary(gls_Wulst_Endo)
#Evomap confidence intervals
Wulst_Endo_CI <- gls.ci(log(Av_data$Wulst_mmsquared),log( Av_data$BrSA_mmsquared-Av_data$Wulst_mmsquared),vcv(consensus_tree))




```




#PGLS results table. Sooo clunkily coded but easier than manually typing in. rbind allows easy addition and swapping (as opposed to hard-coding row numbers).
```{r}

#Set up a table with the relevant column names

PGLS_results <- data.frame("Dep" = integer(0), "Pred"= integer(0), "t"= integer(0), "p"= integer(0))

#Start it off with row 1 

PGLS_results [1,] <-   c( "log Orb. Diam.", "log Geom. Mean", round (summary(gls_orbD)$tTable[2,3:4],3))

PGLS_results <-rbind (PGLS_results, c( "log Opt. For.", "log Geom. Mean",  round (summary(gls_optF)$tTable[2,3:4],3)))

PGLS_results<-  rbind (PGLS_results, c("log Opt. For.", "log Orb. Diam.",round(summary(gls_optF_orbD)$tTable[2,3:4],3)))

PGLS_results <- rbind(PGLS_results, c("log Opt. L. Surf", "log Geom. Mean", round(summary(gls_optLSA)$tTable[2,3:4],3)))

#Then add as many rows with the same outputs as you like using rbind

PGLS_results <-rbind (PGLS_results, c("log Endoc. surf.", "log Geom. Mean", round (summary(gls_EndoLSA)$tTable[2,3:4],3)))

PGLS_results <- rbind (PGLS_results, c( "log Opt. L. Surf.","log Endoc. Surf.", round (summary(gls_optLSA_endo)$tTable[2,3:4],3)))

PGLS_results <- rbind (PGLS_results, c("log Wulst Surf.", "log Endoc.-Wulst surf.", round(summary(gls_Wulst_Endo)$tTable[2,3:4],3)))



#The csv for these is written below with all the normality tests



```



#pANCOVAS - code by Jeroen Smaers

#Do Elanines have larger orbits, smaller Optic foramina for their size; and do they have smaller optic foramina relative to orbit size?
#This code comparesa  model where there is no intercept difference between Elanines with a model where there is one, but the slope is assumed to be the same. Significant p-values mean that the model where Elanines have a different intercept is significantly better.

```{r}


#Orbit diameter vs. geom. mean

Y<-"orbD"; X<-"geomean";

#Note, data logged here
data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")

Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_01_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)

#model is ok

lm_orbD_geomean_Elanus <- lm(Dependent~grpI_01 + Independent,data)


perf_OrbD_Geomean_Elanus <- check_model(lm_orbD_geomean_Elanus )


Sigma<-vcv(tree)
OrbD_geomean_pANOVA <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)


#Optic foramen diameter vs geom mean

Y<-"optF"; X<-"geomean";

data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")

Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_optFGM_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)

lm__optF_geomean_Elanus <- lm(Dependent~grpI_01 + Independent,data)

perf_optF_geomean_Elanus <- check_model(lm__optF_geomean_Elanus )


Sigma<-vcv(tree)
OptF_geomean_pANOVA <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)


#Optic foramen diameter vs orbit diameter

Y<-"orbD"; X<-"optF";

data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")

Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_01_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)


lm_orbD_optF_Elanus <- lm(Dependent~grpI_01 + Independent,data)
perf_OrbD_optF_Elanus <- check_model(lm_orbD_optF_Elanus )


Sigma<-vcv(tree)

OptF_orbD_pANOVA <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)



```


#Do  Elanus (and spotted Harriers, as per plot) have relatively large optic lobe surface areas compared to endocast surface area?


```{r}
#First test for Elanus
#optic lobevs. endocast

Y<-"optLSA"; X<-"BrSA_mmsquared";

data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")


lm_OptL_EndoSA_Elanus <- lm(Dependent~grpI_01 + Independent,data)
perf_OptL_EndoSA_Elanus <- check_model(lm_OptL_EndoSA_Elanus )


Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_01_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)

Sigma<-vcv(tree)
OptL_EndoSA_pANOVA_Elanus <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)


#Pick out the right species, check

tree<-consensus_tree
plot(tree,label.offset = 0.25); tiplabels()
All<-1:length(tree$tip.label)
Circus <- 4

grpI_01<-rep("XXX",length(rownames(data)))
    grpI_01[All]<-"1"
    grpI_01[Circus]<-"2"
    grpI_01<-as.factor(grpI_01); names(grpI_01)<-rownames(data)
    summary(grpI_01); length(grpI_01); length(names(summary(grpI_01)))
    tree$tip.label[which(grpI_01=="XXX")]
        tipCol<-rep("black",length(tree$tip.label))
        tipCol[which(grpI_01=="1")]<-"red"
        tipCol[which(grpI_01=="2")]<-"darkgreen"
        plot(tree,tip.color = tipCol)
    
#Optic lobe vs geomean

Y<-"optLSA"; X<-"geomean";

data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")

lm_optLSA_geomean_Circus <- lm(Dependent~grpI_01 + Independent,data)
perf_optLSA_geomean_Circus <- check_model(lm_optLSA_geomean_Circus)

Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_01_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)

Sigma<-vcv(tree)
OptL_geomean_pANOVA_Circus <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)

#Optic lobe surface vs. endocast surface


Y<-"optLSA"; X<-"BrSA_mmsquared";

data<-log(na.omit(Av_data[,c(which(colnames(Av_data)==paste(Y)),which(colnames(Av_data)==paste(X))),drop=F]))

colnames(data)<-c("Dependent","Independent")

lm_optLSA_EndoSA_Circus <- lm(Dependent~grpI_01 + Independent,data)
perf_optLSA_EndoSA_Circus <- check_model(lm_optLSA_EndoSA_Circus)

Model<-model.matrix(as.formula(Dependent~Independent),data)
Model_01_I<-model.matrix(as.formula(Dependent~grpI_01 + Independent),data)

Sigma<-vcv(tree)
OptL_EndoSA_pANOVA_Circus <- gls.ancova(Dependent~Independent,Sigma,Model,Model_01_I)



pANOVA_summary <- rbind(OrbD_geomean_pANOVA, OptF_geomean_pANOVA, OptF_orbD_pANOVA, OptL_EndoSA_pANOVA_Elanus,  OptL_EndoSA_pANOVA_Circus,OptL_geomean_pANOVA_Circus)



write.csv(pANOVA_summary, file = "../Tables/pANOVA_output.csv")





```



#Some checks for model performance

```{r}

#A pdf with model linearity, collinearity,and residual normality ; not reported, but just for pretty checks.

# pdf(file = "../Figures/performance_plots.pdf", height = 10, width = 8, onefile = TRUE)
# 
# perf_optLSA
# perf_EndoLSA
# perf_optLSA_endo
# 
# perf_orbD
# perf_optF
# perf_optF_orbD
# 
# perf_OrbD_Geomean_Elanus
# perf_optF_geomean_Elanus
# perf_OrbD_optF_Elanus
# 
# perf_OptL_EndoSA_Elanus
# 
# 
# dev.off()


normchecks <- c("lm_orbD","lm_optF", "lm_optF_orbD","lm_optLSA","lm_EndoLSA","lm_optLSA_endo","lm_Wulst_Endo","lm_orbD_geomean_Elanus", "lm__optF_geomean_Elanus" ,"lm_orbD_optF_Elanus","lm_OptL_EndoSA_Elanus", "lm_optLSA_EndoSA_Circus","lm_optLSA_EndoSA_Circus","lm_optLSA_geomean_Circus"  )

model_res_norm <- matrix(nrow = length(normchecks), ncol = 2)

for (i in c(1: length(normchecks))){
  
  
  temp <- check_normality(get(normchecks[i]))
  
  model_res_norm [i,1] <- paste (normchecks[i])
  model_res_norm[i,2] <- round( temp[1], digits = 3)
  
  
}


#Just for pasting pgls results, the PANOVA results need to taken manually from model_res_norm

PGLS_results <- cbind(PGLS_results, model_res_norm[c(1:nrow(PGLS_results)),2])


write.csv(PGLS_results, file = "../Tables/PGLS_output.csv")

```


#Are Elanus species, Circus, or Aviceda, significantly different in orbit convergence?

```{r}
Elanus_gp <- rep(1, length(rownames(Av_data)))
Elanus_gp [which (Av_data$Genus=="Elanus")] <- 2

Circus_gp <- rep(1, length(rownames(Av_data)))
Circus_gp [which (Av_data$Genus=="Circus")] <- 2

Aviceda_gp <- rep(1, length(rownames(Av_data)))
Aviceda_gp [which (Av_data$Genus=="Aviceda")] <- 2

#the phylANOVAs give a warning message that there are no labels for x, assuming order of tree$tip.label. Not sure why but the tips and data are in the same order:

rownames(Av_data) == tree$tip.label

Deg_Elanus <- phylANOVA(tree, Elanus_gp, Av_data$OrbP_Degrees,nsim=10000)

Deg_Circus <- phylANOVA(tree, Circus_gp, Av_data$OrbP_Degrees,nsim=10000)

Deg_Aviceda <- phylANOVA(tree, Aviceda_gp, Av_data$OrbP_Degrees,nsim=10000)

Degree_phylANOVA <- 
rbind(
c("Genus",  "MeanSq", "F", "p"),
c ("Elanus",  Deg_Elanus$`Mean Sq`[1], Deg_Elanus$F, Deg_Elanus$Pf),
c ("Circus", Deg_Circus$`Mean Sq`[1], Deg_Circus$F, Deg_Circus$Pf),
c ("Aviceda", Deg_Aviceda$`Mean Sq`[1], Deg_Aviceda$F, Deg_Aviceda$Pf)
)

write.csv(Degree_phylANOVA, file = "../Tables/Degree_phylanova.csv")


```


##############Plots- set up

```{r, echo=TRUE}
#pdf("../Figures/Scatterplot_Geomean.pdf", height = 3, width = 9)


colsSpecies=c("red3","firebrick4", "indianred1", "forestgreen", "hotpink",  "blue", "darkblue","darkslateblue", "dark orange", "darkred", "wheat3","deepskyblue", "azure4", "skyblue", "cyan",   "blueviolet")

#adding some plotting parameters that mean you can change them all at once for each subplot. You can also specify colours this way
lwd_abline <- 2
lwd_CI <- 2
lty_CI <- 2
col_CI <- "hotpink"

names(colsSpecies)<- levels (as.factor(Av_data$GenSpec))
pchPLOT=c(19,8,8, 17, 15)
cexes = c(2,2.5,2.5,2,2)
Av_data$desig <- rep(1,length(rownames(Av_data))); Av_data$desig[which(Av_data$species=="scriptus")] <- 2;
Av_data$desig[which(Av_data$species=="axillaris")] <- 3;

Av_data$desig[which(Av_data$species=="caeruleus")] <- 3



Full_data$desig <- rep(1,length(rownames(Full_data)));
Full_data$desig[c(which(Full_data$species=="scriptus"),which(Full_data$species=="axillaris"), which(Full_data$species == "caeruleus"))] <- 2
Full_data$desig[which(Full_data$species=="subcristata")] <- 4;
Full_data$desig[which(Full_data$species=="assimilis")] <- 5



```


Phylogeny

```{r}

pdf(file="../Figures/Phylogeny.pdf")

plot(consensus_tree, tip.color = colsSpecies [Av_data$GenSpec])

dev.off()


```



#This is plotting the evomap CI intervals based confidence intervals. Note that I am using upper/lower 2.5% of the distribution (a total of 5% because we're looking at both tails).


#Orbit/optic foramen plots. Really lazy in double-plotting but whatever ;-); saves me setting up a complicated line of plot just to prepare


```{r}

pdf("../Figures/Orbit_optic.pdf", height = 10, width = 5)

par(mfrow=c(4,1), bty="n", mar = c(rep(4,4)))

xlim1 <-  c(2.24,2.82)  
ylim_orb_geo <- c(2.65,3.4)
ylim_opt_geo <- c(0.8,1.5)
ylim_opt_orb <- c(0.8,1.5)

cexesOrbopt<-cexes-0.6

#Orbit diameter against geomean
plot(log (Full_data$orbD)~ log (Full_data$geomean),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Orbit Diameter (mm)", xlab="log Geom. Mean", xlim=xlim1, ylim = ylim_orb_geo, bty="o")


#Here we make the grey PI polygon. To make that appear behind the plot, I have to repeat the plot command. Not elegant but works!

polygon(c(OrbD_gm_PI$PI.plot$X, rev (OrbD_gm_PI$PI.plot$X)), c(OrbD_gm_PI$PI.plot$Lower2.5, rev(OrbD_gm_PI$PI.plot$Upper2.5)), col = grey (0.9), border = NA )

par(new=TRUE)

plot(log (Full_data$orbD)~ log (Full_data$geomean),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Orbit Diameter (mm)", xlab="log Geom. Mean", xlim=xlim1, ylim = ylim_orb_geo,bty="o");

lines(OrbD_gm_CI$CI.plot$X,OrbD_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
lines(OrbD_gm_CI$CI.plot$X,OrbD_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)

  #I am here using the abline from the evomap code, but the pgls with brownian settings results in teh same line
abline (OrbD_gm_CI$model[1,1], OrbD_gm_CI$model[2,1])

mtext(text="A",side=1,line=-10,outer=FALSE, adj = 0.04,cex=0.7)



#Optic foramen diameter against geomean
plot(log (Full_data$optF)~ log (Full_data$geomean),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Optic Foramen Diameter (mm)", xlab="log Geom. Mean",xlim=xlim1, ylim=ylim_opt_geo,bty="o");
text((Full_data$optF)~(Full_data$geomean), labels='', adj=1, cex=0.7)


polygon(c(OptF_gm_PI$PI.plot$X, rev (OptF_gm_PI$PI.plot$X)), c(OptF_gm_PI$PI.plot$Lower2.5, rev(OptF_gm_PI$PI.plot$Upper2.5)), col = grey (0.9), border = NA )

par(new=TRUE)


plot(log (Full_data$optF)~ log (Full_data$geomean),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Optic Foramen Diameter (mm)", xlab="log Geom. Mean",xlim=xlim1, ylim=ylim_opt_geo,bty="o");
text((Full_data$optF)~(Full_data$geomean), labels='', adj=1, cex=0.7)



abline (OptF_gm_CI$model[1,1], OptF_gm_CI$model[2,1],lwd = lwd_abline)
#The coefficients from the pgls are [1] intercept, [2] slope, and you can fit this to the graph like yay - there are cosmetics you can look up by typing ?abline.
#abline( gls_optF$coefficients[1],gls_optF$coefficients[2], lwd = lwd_abline  )



lines(OptF_gm_CI$CI.plot$X,OptF_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
lines(OptF_gm_CI$CI.plot$X,OptF_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)
  
mtext(text="B",side=1,line=-10,outer=FALSE, adj = 0.04, cex = 0.7)

  
  
     
#Orbit diameter against optic Foramen diameter
  
plot(log (Full_data$optF)~ log (Full_data$orbD),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Optic Foramen Diameter (mm)", xlab="log Orbit Diameter (mm)",ylim= ylim_opt_orb, bty="o");
text((Full_data$orbD)~(Full_data$optF),labels='', adj=1, cex=0.7)


polygon(c(OptF_OrbD_gm_PI$PI.plot$X, rev (OptF_OrbD_gm_PI$PI.plot$X)), c(OptF_OrbD_gm_PI$PI.plot$Lower2.5, rev(OptF_OrbD_gm_PI$PI.plot$Upper2.5)), col = grey (0.9), border = NA )

par(new=TRUE)

plot(log (Full_data$optF)~ log (Full_data$orbD),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexesOrbopt[Full_data$desig], ylab="log Optic Foramen Diameter (mm)", xlab="log Orbit Diameter (mm)",ylim= ylim_opt_orb, bty="o");
text((Full_data$orbD)~(Full_data$optF),labels='', adj=1, cex=0.7)

abline( gls_optF_orbD$coefficients[1],gls_optF_orbD$coefficients[2], lwd = lwd_abline  )

lines(OptF_OrbD_gm_CI$CI.plot$X,OptF_OrbD_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
lines(OptF_OrbD_gm_CI$CI.plot$X,OptF_OrbD_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)


mtext(text="C",side=1,line=-10,outer=FALSE, adj = 0.04, cex=0.7)
  

plot(1, type = "n",                         # Remove all elements of plot
     xlim = c(0, 10), ylim = c(0, 10),
     axes=F,
     xlab=NA, ylab=NA
     
     )
legend(-0.6,11 , levels(as.factor(Full_data$GenSpec)) , text.font = 3, pch= c(19,19,19,17,15,8,8,8,rep(19,8)), col=colsSpecies[levels(as.factor(Full_data$GenSpec))], cex=1.1, pt.cex=1.4, bty="n",ncol=3 )




dev.off()
  
```

#Endocast surface/size, optic lobe/size,

```{r}

pdf("../Figures/Surfaces.pdf", height = 10, width = 10)

#The layout and legend plotting is inspired by code by Emma Sherratt :-) in Weisbecker et al. 2021.

#This tells it to make 3 rows and (therefore 2 columns) with images 1-4 each occupying one position and image 5 occupying 2 positions. Thanks to Emma Sherratt for this trick
mat <- matrix(c(1,2,3,4,5,5), nrow=3,byrow=T)

# set the size of fields
layout(mat, widths=c(1,1,0.5,0.5), heights=c(0.3,0.3,0.12))
#Plot PC1 vs PC2
par(mar=c(4, 5, 0.5, 1), cex.lab = 1.5, cex.axis =1.5)  # sets the margins c(bottom, left, top, right)



  
#Optic lobe against geomean
plot(log(Full_data$optLSA)~log(Full_data$geomean),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexes[Full_data$desig], ylab="log Av. Optic Lobe Surf. Area (mm2)", xlab=" log Geomean");


#The coefficients from the pgls are [1] intercept, [2] slope, and you can fit this to the graph like yay - there are cosmetics you can look up by typing ?abline.
abline( gls_optLSA$coefficients[1],gls_optLSA$coefficients[2], lwd = lwd_abline  )

  lines(OptLSA_gm_CI$CI.plot$X,OptLSA_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
  lines(OptLSA_gm_CI$CI.plot$X,OptLSA_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)

mtext(text="A",side=1,line=-25,outer=FALSE, adj = 0.04, cex=0.9)

#Endocast surf. lobe against geomean
plot(log (Full_data$BrSA_mmsquared) ~ log (Full_data$geomean),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexes[Full_data$desig], ylab= "log Endoc. Surf. Area (mm2)", xlab="log Geomean");

abline( gls_EndoLSA$coefficients[1],gls_EndoLSA$coefficients[2], lwd = lwd_abline  )

  lines(EndoLSA_gm_CI$CI.plot$X,EndoLSA_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
  lines(EndoLSA_gm_CI$CI.plot$X,EndoLSA_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)
  
mtext(text="B",side=1,line=-25,outer=FALSE, adj = 0.04, cex=0.9)  
  
  #Optic lobe against endocast SA - optic lobe
plot(log (Full_data$optLSA)~ log ( Full_data$BrSA_mmsquared - Full_data$optLSA ),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexes[Full_data$desig], ylab="log Optic Lobe Surf. Area (mm2)", xlab="log Endoc. surf. Area minus Optic Lobe Surf. Area (mm2)");


abline( gls_optLSA_endo$coefficients[1],gls_optLSA_endo$coefficients[2], lwd = lwd_abline  )

  lines(OptLSA_Endo_gm_CI$CI.plot$X,OptLSA_Endo_gm_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
  lines(OptLSA_Endo_gm_CI$CI.plot$X,OptLSA_Endo_gm_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)  
  
 mtext(text="C",side=1,line=-25,outer=FALSE, adj = 0.04, cex=0.9) 
  
#Wulst vs endocast

plot(log(Full_data$Wulst_mmsquared)~log(Full_data$BrSA_mmsquared - Full_data$Wulst_mmsquared),col=colsSpecies[Full_data$GenSpec], pch=pchPLOT[Full_data$desig], cex=cexes[Full_data$desig], ylab=" log Wulst Surf. Area (mm sq.)", xlab="log Endocast Surf. minus Wulst surf.");


#The coefficients from the pgls are [1] intercept, [2] slope, and you can fit this to the graph like yay - there are cosmetics you can look up by typing ?abline.
abline( gls_Wulst_Endo$coefficients[1],gls_Wulst_Endo$coefficients[2], lwd = lwd_abline  )

  lines(Wulst_Endo_CI$CI.plot$X,Wulst_Endo_CI$CI.plot$Upper2.5,col= col_CI, lwd= lwd_CI, lty= lty_CI)
  lines(Wulst_Endo_CI$CI.plot$X,Wulst_Endo_CI$CI.plot$Lower2.5, col= col_CI, lwd=lwd_CI, lty = lty_CI)  
  
  
mtext(text="D",side=1,line=-25,outer=FALSE, adj = 0.04, cex=0.9)  
  
  #Make new empty plot for legend
  
  plot(1, type = "n",                         # Remove all elements of plot
     xlim = c(0, 10), ylim = c(0, 10),
     axes=F,
     xlab=NA, ylab=NA
     
     )
legend(1,10 , levels(as.factor(Full_data$GenSpec)) , text.font = 3, pch= c(19,19,19,17,15,8,8,8,rep(19,8)), col=colsSpecies[levels(as.factor(Full_data$GenSpec))], cex=1.5, pt.cex=2.1, bty="n",ncol=4, y.intersp = 1.2 )
  
dev.off()
  
```

#Orbit posterior angle single graph
```{r}
pdf("../Figures/Orbit_Posterior_Angle_AllSpecies.pdf")
par(mfrow=c(1,1.5), bty="O")


a <- rep(1, nrow(Full_data))



plot(0:50,0:50, xlim=c(0.5,2),ylim=c(35,85), type="n", xaxt="n",
     xlab = "",
     ylab = "Degrees"  )

axis(1, at=c(1), labels=c("Orbit Posterior Angle"));

points(y=Full_data$OrbP_Degrees, x=(a),col=colsSpecies[Full_data$GenSpec], cex = cexes[Full_data$desig]-0.1, pch=pchPLOT[Full_data$desig]);


legend(1.5,70 , levels(as.factor(Full_data$GenSpec)) , pch= c(19,19,19,17,15,8,8,8,rep(19,8)), col=colsSpecies[levels(as.factor(Full_data$GenSpec))], cex=1, bty="n", pt.cex=1.5 , text.font = 3)

dev.off()

```

###This graph was removed from the manuscript and replaced with the graph above. It included data which added confusion and was irrelevant, thus it was taken out but is still included below.
#Multiple degree measures on one graph - Note that the legend is shoddily coded right now but that's just more convenient for now.
```{r}
pdf("../Figures/Orbit_Angle_AllSpecies.pdf")
par(mfrow=c(1,1), bty="L")


a <- rep(1, nrow(Full_data))
b <- rep(2, nrow(Full_data))
c <- rep(3, nrow(Full_data))


plot(0:50,0:50, xlim=c(0.5,5),ylim=c(10,80), type="n", xaxt="n",
     xlab = "",
     ylab = "Degrees"  )

axis(1, at=c(1,2,3), labels=c("Orb. Post.","Orb. Ant.", "Orb. Post./Sup."));

points(y=Full_data$OrbP_Degrees, x=(a),col=colsSpecies[Full_data$GenSpec], cex = cexes[Full_data$desig]-0.3, pch=pchPLOT[Full_data$desig]);

points(y=Full_data$OrbS_Degrees, x=(b),col=colsSpecies[Full_data$GenSpec], cex = cexes[Full_data$desig]-0.3,pch=pchPLOT[Full_data$desig]);

points(y=Full_data$orbPS, x=(c),col=colsSpecies[Full_data$GenSpec], cex = cexes[Full_data$desig]-0.3,pch=pchPLOT[Full_data$desig]);

legend(3.7,60 , levels(as.factor(Full_data$GenSpec)) , pch= c(19,19,19,17,15,8,8,8,rep(19,8)), col=colsSpecies[levels(as.factor(Full_data$GenSpec))], cex=1, bty="n", pt.cex=1.5 , text.font = 3)

dev.off()

```
