---
title: "Data and phylogeny processing for: Not like night and day: the nocturnal Letter-winged Kite does not differ from diurnal congeners in orbit or endocast morphology"
author: "Vera Weisbecker, Jeroen Smaers,Aubrey Keirnan, and Andrew Iwaniuk"
date: "February 3rd, 2022"
output: html_document
---

#load phylogenetic packages

```{r}

library(ape)
library(geiger)
library(nlme)

library(rstudioapi)

#this sets the working directory to the home path, which doesn't happen on everyone's computers
setwd(dirname(getActiveDocumentContext()$path))


```

#Reading and processing raw data

```{r}
#Read data
Data_raw <- read.csv(file="../Data/Raw/Raptor_Data_Full.csv", header=TRUE)
#Give each specimen its number to get unique row names
rownames(Data_raw) <- c(paste(substr(Data_raw$Genus,1,1),substr(Data_raw$species,1,4), "_", Data_raw$Museum, Data_raw$Catalogue_number ,sep = ""))

#This is the only case where there are 2 NAs (in M. sphenurus and A. subcristata). In these 2 cases, entering the value just for the left:
colSums(is.na(Data_raw))

Data_raw$OptFD_L_mm[which(is.na(Data_raw$OptFD_L_mm)==TRUE)] <- Data_raw$OptFD_R_mm[which(is.na(Data_raw$OptFD_L_mm)==TRUE)]

colSums(is.na(Data_raw))

#averaging lefts and rights - note na not required to be considered because dataset has both sides in all cases. Note the Wulst data are the sum of both areas from the outset (no left and right).
optLSA <- rowMeans(data.frame(Data_raw$OptLSA_R_mmsquared,Data_raw$OptLSA_L_mmsquared)[,1:2])
optF <- rowMeans(data.frame(Data_raw$OptFD_R_mm,Data_raw$OptFD_L_mm)[,1:2])
orbD <- rowMeans(data.frame(Data_raw$OrbitD_R_mm,Data_raw$OrbitD_L_mm)[,1:2])
orbPS <- rowMeans(data.frame(Data_raw$OrbPS_R_Degrees,Data_raw$OrbPS_L_Degrees)[,1:2])

data_lr <- cbind(Data_raw, optLSA, optF, orbD, orbPS)


```

#Adding geometric mean as size proxy and putting together in an ordered fashion

```{r, echo=TRUE}

#Specimen provenance
Specimen_provenance_data <- Data_raw[,c(1:5)]

#Data that might be allometric
Size_containing_data <- cbind (Data_raw[,c(6:16,25:27)], optLSA, optF, orbD, orbPS)

#getting the geomean only from the data without left and right averaged. Making it into a column. Also only including linear measurements, not surface and volumes because they likely change between species according to different scales


geomean <- apply(Size_containing_data[,c(4,8:14)], 1, prod)^(1/ncol(Size_containing_data[,c(4,8:14)]))

#Data of degrees; these are limited to 360 degrees and not expected to be related to size 
Degrees_data <- Data_raw[ ,c(17:24)]

#Data that contain other factors and are non-numeric

Factorial_data <- Data_raw[ ,c(28:32)]


#A full dataset with all the different data types one after the other
Full_data <- cbind(Specimen_provenance_data, Size_containing_data, geomean,Degrees_data,Factorial_data)
rownames(Full_data) <- rownames(Data_raw)

Full_data$GenSpec <- paste(substr( Full_data$Genus, start=1, stop=1), ". ", Full_data$species, sep="")

```


#Averaging species
```{r}
#averaging species 
#data_av <- cbind (aggregate(Full_data_plus_resids[,c(6:29)], by = list(Full_data_plus_resids$species), FUN = mean), aggregate(Full_data_plus_resids[, c (2:4, 46, 48, 49)], by = list(Full_data_plus_resids$species), FUN = unique))


Av_data <- cbind (aggregate((Full_data[,c(8:32)]), by = list(Full_data$species), FUN = mean), aggregate(Full_data[, c (1:3, 33,  36,37)], by = list(Full_data$species), FUN = unique))

#We don't want the new Group.1 column. No idea where taht comes from or the purpose!
Av_data <- Av_data [,-which (colnames(Av_data)=="Group.1",),]

#the row names of this should be full names so they can be matched with the tree - but here we make an abbreviation for plotting
Av_data$Abbreviation <- c(paste(substr(Av_data$Genus,1,1),substr(Av_data$species,1,4), sep = ""))

#row names
row.names(Av_data) <-c(paste(Av_data$Genus, Av_data$species, sep="_"))


head(Av_data)

Av_data$GenSpec <- paste(substr( Av_data$Genus, start=1, stop=1), ". ", Av_data$species, sep="")

#VW conducted quick visual spot check in 3 random positions with original dataset that all is properly averaged. Can't be too careful! All good.

```

#load and create phylogeny - hashed out because not used again

```{r}

#Getting the tree from tree_pruner 
# 
 Trees <- read.nexus("../Data/Raw/Phylogeny/tree-pruner-fbbae0fa-904c-4174-b7e3-b81f992e8264/output.nex")
# 
# #Creating majority consensus tree
# 
 Interim_constree_95 <- consensus(Trees, p=0.95)
 
 #Using the below consensus tree because it resolves Elanus better and in keeping with other phylogenies
 Interim_constree_50 <- consensus(Trees, p=0.50)
# 
# write.nexus(Interim_constree, file = "../Data/Raw/Phylogeny/Consensus_tree.nex")
```

#This file is then manually changed in Mesquite so that _Circus_ is outside of _Accipiter and _E. caeruleus_ is outside the other two Elanus (following a combinaiton of Minden et al. 2018 and Starikove and Wink 2020 ) NOTE also I made this tree ultrametric. This means we only look at the topology, not divergence times. This is necessary because the Circus and Elanus positions have no time on it due to the manual rearrangement. 

```{r}

consensus_tree <- compute.brlen( read.nexus("../Data/Raw/Phylogeny/Consensus_tree_Circus_Elanus_resolved.nex"))

plot(consensus_tree)

name.check(consensus_tree, Av_data)

#Reorder the average dataset so the row names match up 

Av_data  <- Av_data[match( consensus_tree$tip.label, rownames(Av_data)),]

Av_data$Species <- rownames(Av_data)

#Check that this is correct

rownames(Av_data) == consensus_tree$tip.label

```


#Save everything as an rdata file so there's no need to call everything singly
```{r}

save(Data_raw,Full_data,Av_data, consensus_tree, file = "../Data/Processed/LWK_data_processed.rda")


```
